---
author: "Anders Torstensson"
date: "`r Sys.Date()`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile,
                    encoding = encoding, 
                    output_file = paste0("output/html_reports/ifcb_data_export_report_",
                                         format(Sys.time(), "D%Y%m%dT%H%M"),
                                         ".html"))
})
title: "`r paste('ifcb-pipeline:', params$classifier, 'classifier')`"
---

## IFCB Data Pipeline Report

This is a dynamically generated report based on classifier and year data passed through the `render()` function.

### Classifier Information
```{r}
# Define default values if params object does not exist
if (!exists("params")) {
  params <- list(
    classifier = "Skagerrak-Kattegat",
    year = c(2022, 2023, 2024),
    threshold = "opt",
    remove_flagged_data = c("bubbles", "incomplete", "near land"),
    regional = TRUE,
    multiyear_delivery = TRUE,
    f1_threshold = 0.9
  )
}

# Print classifier and year info
cat("Classifier: ", params$classifier, "\n")
cat("Years: ", paste(params$year, collapse = ", "), "\n")
cat("Threshold: ", params$threshold, "\n")
cat("Remove flagged data: ", paste(params$remove_flagged_data, collapse = ", "), "\n")
cat("Multiyear delivery: ", params$multiyear_delivery, "\n")
cat("F1 Threshold: ", params$f1_threshold, "\n")
```
### Data Processing Summary

```{r}
# Simulate some data processing summary
data_summary <- data.frame(
  Year = params$year,
  Processed_Items = sample(1000:5000, length(params$year)),
  Flagged_Items = sample(50:300, length(params$year))
)

# Print data summary
knitr::kable(data_summary, caption = "Summary of Processed Data")
```

### Additional Information

This report was generated dynamically based on parameters passed to the document. You can adjust the parameters using the `render()` function in the R script.
